---
title: "Regression - Statistical Analysis"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(patchwork)
library(readxl)
library(kableExtra)
```


```{r echo=FALSE} 
original = tibble(        ## Clean the original weight data
  read.csv("./dataset/Student_Weight_Status_Category_Reporting_Results__Beginning_2010.csv")
) %>%
  janitor::clean_names() %>%
  select(-location_code, -region, -area_name)  # the only location information we need is county name
```


```{r echo=FALSE} 
# Clean data set with geo location information
coordinates = tibble(
  read.csv("./dataset/Geocodes_USA_with_Counties.csv")
) %>%
  filter(state == "NY") %>%  # filter out counties outside NY state
  select(county, latitude, longitude) %>% # only information we need is county name and geo location
  drop_na() %>%
  group_by(county) %>%
  summarise(latitude = mean(latitude), longitude = mean(longitude)) %>% #different location in each county varied slightly, so we take the mean of each county's geo location
  filter(!county == "") %>% # one county's name input is blank
  mutate(county = toupper(county)) # to switch county name to uppercase
```


```{r echo=FALSE}
# combine two data set
weight_df = left_join(original, coordinates, by = "county")
```


```{r echo=FALSE}
# Further original data cleaning for linear regression analysis
linear_df = weight_df %>%
  filter(!sex == "ALL") %>%
  filter(!grade_level == "DISTRICT TOTAL") %>%
  mutate(
    sex = if_else(sex == "MALE", 0, 1),
    grade_level = if_else(grade_level == "ELEMENTARY", 0, 1)
  ) %>%
  drop_na() %>%
  filter(year_reported %in% c("2018-2019")) %>% #only analyzing data in year 2018-2019
  select(county, percent_overweight_or_obese, grade_level, sex)
```


```{r echo=FALSE, warning=FALSE}
# Import tidy and join the median income and food insecurity and race data
income = read_xlsx("./dataset/median_income.xlsx") %>%
  janitor::clean_names() %>%
  rename(county = region_county) %>%
  mutate(median_income = median_income*0.001,  # convert the income unit from $100,000 to 100k format, large values will reduce model's efficiency
         county = toupper(county))

food_insecurity = read_xlsx("./dataset/food_insecurity.xlsx") %>%
  janitor::clean_names() %>%
    rename(county = region_county) %>%
  rename(food_insecurity_p = percentage) %>%
  mutate(county = toupper(county))


linear_df2 = left_join(linear_df, income, by = "county")  # combine income data to the main 
linear_df3 = left_join(linear_df2, food_insecurity, by = "county") # combine food insecurity data
```

# Normality Check and Data Transformation

Before fitting the linear regression model, I used two types of transformation to improve model adequacy, a. y' = log(y), b. y'= y^0.5, Q_Q plot showed that the log tansformation has improved data's normality.  

```{r echo=FALSE, fig.align='center'}
# Normality check
qqnorm(log(linear_df$percent_overweight_or_obese), col = c("darkorchid3"))

linear_df3 %>%
  ggplot(aes(x = log(percent_overweight_or_obese))) + geom_histogram(color = "purple",
    fill = "#69b3a2", size = 2) + labs(x = "log (percent overweight and obese)", title = "Histogram of log (percent overweight and obese)") 
```

Above are the Q-Q plot and histogram of overweight/obesity data after log transformation.  Q-Q plot is in a perfect straight line, histogram is normally distributed but a bit skewed to the left.  Overall, the normality assumption is satisfied.  

Based on the transformation above, the model we are going to fit is:
$$ log(percentage \space overweight \space or \space obese) = \beta_0 + \beta_1grade \space level + \beta_2median \space income + \beta_3 food \space insecurity \space rate + \beta_4 gender$$

# Statistics Model

In the beginning we included 4 independent variables, grade level, median income, food insecurity,  gender, however the model summary showed there is a weak correlation between gender (p-value = 0.168) with dependent variable y, so we took gender out from the model, left three variables all significantly influenced the percentage overweight and obese.

# Final Model: 
$$ log(y) = 4.541 + 0.161x_1 - 0.010x_2 - 0.039x_3$$
y = percentage overweight or obese  
x1= grade level  (0 = elementary, 1 = middle/high school)  
x2 = median income (k dollars)  
x3 = food insecurity percentage

```{r echo=FALSE, warning=FALSE}
# regression model
lm_1 = lm(log(percent_overweight_or_obese) ~ grade_level  + median_income + food_insecurity_p, data = linear_df3) 
```
# Model Summary
```{r}
summary(lm_1)
```

```{r echo=FALSE, warning=FALSE, fig.align='center'}
par(mfrow = c(2, 2))
plot(lm_1, col = "darkorchid2", alpha = 0.5)
```

# Conclusion
There is no strange pattern about Residusl vs Fitted plot, and Normal Q-Q plot is on a straight line, so the constant variance assumption is met, and this model is valid.  Three variables are significant with very small p-value (<0.001).  However, due to the limitation of data, the model's R-squared is equal to 0.2526, which means only 25% of the data is explained by this model.  A possible explanation is that other important factors have not been included.  We will make improvement when more data is available. 
